### 2019-01-31 abstract.tex
ssr

A class of Boolean functions, 
called \textbf{nested canalyzing functions} (NCFs),
has been used to model certain biological phenomena.
Researchers have studied some properties of these functions
such as sensitivity and stability.
In this note, we examine the relationships between NCFs and symmetric 
Boolean functions. 
We identify all the Boolean functions that NCFs as well as symmetric. 
A more general form of symmetric Boolean functions,
called $r$-symmetric functions (where $r$ is the symmetry level),
has also been studied in the literature.
%% We establish lower and upper bounds on the symmetry
%% level of an NCF $f$ as functions of the number of layers
%% in a standard respresentation of $f$.  
We observe that, in general,
one cannot efficiently approximate the symmetry level of
a Boolean function to within any factor, unless \textbf{P} = \cnp. 
In contrast, we show that the symmetry level of an NCF $f$
can be easily computed from a standard (and efficiently
computable) representation of $f$.

>>>>

A class of Boolean functions, 
called \textbf{nested canalyzing functions} (NCFs),
has been used to model certain biological phenomena.
Researchers have studied some properties of these functions
such as sensitivity and stability.
In this note, we examine the relationships between NCFs and symmetric 
Boolean functions. 
We identify all the Boolean functions that NCFs as well as symmetric. 
A more general form of symmetric Boolean functions,
called $r$-symmetric functions (where $r$ is the symmetry level),
has also been studied in the literature.
We observe that, in general,
one cannot efficiently approximate the symmetry level of
a Boolean function to within any factor, unless \textbf{P} = \cnp. 
In contrast, we show that the symmetry level of an NCF $f$
can be easily computed given a standard representation of $f$.
We also that for any NCF with $n$ variables, the notion of
strong asymmetry considered in the literature is equivalent to
the property that the symmetry level of the function is $n$.

### 2019-01-31 intro.tex
ssr

>>>>

Other forms of symmetry which can be more general than the permutations 
corresponding to symmetry groups have been considered \cite{KS-2000}.
We say that a Boolean function $f(x_1, x_2, \ldots, x_n)$
is \textbf{strongly asymmetric}
if for any permutation $\pi$ of $\{1, 2, \ldots, n\}$
\emph{except} the identity permutation,
there exists an input $(a_1, a_2, \ldots,  a_n)$
to $f$ such that $f(a_1, a_2, \ldots, a_n)$ $\neq$
$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
In general, there are functions with $n$ variables that are properly
$n$-symmetric, but not strongly asymmetric \cite{KS-2000}.
However, in the case of NCFs with $n$ variables, we show that the notion of 
strong asymmetry coincides with that of being properly $n$-symmetric.

### 2019-01-31 symmetry_properties.tex
ssr

\subsection{Overview}
\label{sse:res_overview}

@@

\begin{theorem}\label{thm:ncf_symmetric}
The only symmetric canalyzing functions are OR, AND, NOR, NAND, 
the constant function 0, and the constant function 1.

The only symmetric NCFs are OR, AND, NOR, and NAND.
\end{theorem}
\noindent
\textbf{Proof:}~
Suppose symmetric function $f$ is a canalyzing function. Then there
is a variable $x$, and values $a$ and $b$ such that whenever $x =
a$, function $f$ has value $b$.  Since $f$ is symmetric, $f$  has
value $b$ whenever any of its variables have value $a$.  Thus, if
at least one of its variables has value $a$, then $f$ has value
$b$.  If $f$ has value $\bar{b}$ when none of its variables has
value $a$, the four possible combinations of values for $a$ and $b$
correspond to the four functions OR, AND, NOR, and NAND.  If $f$
has value $b$ when none of its variables has value $a$, then $f$
is the constant function $b$.

@@

\subsection{Symmetric Pairs of Variables in an NCF: A Characterization
and Its Applictions}
\label{sse:ncf_strong_sym}

@@

Generalized forms of symmetry have been considered \cite{KS-2000}
in which a Boolean function is symmetric with respect to some permutation,
which can be more general than the permutations corresponding to symmetry groups.
We say that a Boolean function $f(x_1, x_2, \ldots, x_n)$ 
is \textbf{strongly asymmetric}
if for any permutation $\pi$ of $\{1, 2, \ldots, n\}$
\emph{except} the identity permutation,
there exists an input $(a_1, a_2, \ldots,  a_n)$ 
to $f$ such that $f(a_1, a_2, \ldots, a_n)$ $\neq$  
$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
In general, there are functions with $n$ variables that are properly
$n$-symmetric, but not strongly asymmetric \cite{KS-2000}.

We now show that any $n$ variable NCF that is properly $n$-symmetric
is also strongly asymmetric.
As a consequence, there are NCFs with $n$ variables
which are properly $n$-symmetric.

>>>>

\subsection{Overview}
\label{sse:res_overview}
Our first result (Proposition~\ref{pro:ncf_symmetric})
identifies all the functions that are symmetric and
canalyzing as well as those that are symmetric and NCFs.
Our next result (Theorem~\ref{thm:ncf_symmetric_variables})
provides a characterization of pairs of variables of an NCF
that are part of the same symmetry group. 
This characterization allows us to give a simple closed form
expression (Theorem~\ref{thm:ncf_r_symmetric}) 
for the symmetry level of an NCF specified using
its normalized representation.

@@

\begin{proposition}\label{pro:ncf_symmetric}
The only symmetric canalyzing functions are OR, AND, NOR, NAND, 
the constant function 0, and the constant function 1.

The only symmetric NCFs are OR, AND, NOR, and NAND.
\end{proposition}
\noindent
\textbf{Proof:}~
Suppose symmetric function $f$ is a canalyzing function. Then there
is a variable $x$, and values $a$ and $b$ such that whenever $x =
a$, function $f$ has value $b$.  Since $f$ is symmetric, $f$  has
value $b$ whenever any of its variables have value $a$.  Thus, if
at least one of its variables has value $a$, then $f$ has value
$b$.  If $f$ has value $\bar{b}$ when none of its variables has
value $a$, the four possible combinations of values for $a$ and $b$
correspond to the four functions OR, AND, NOR, and NAND.  If $f$
has value $b$ when none of its variables has value $a$, then $f$
is the constant function $b$.

@@

\subsection{Symmetric Pairs of Variables in an NCF: A Characterization
and Its Applications}
\label{sse:ncf_strong_sym}

@@

\iffalse
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Generalized forms of symmetry have been considered \cite{KS-2000}
in which a Boolean function is symmetric with respect to some permutation,
which can be more general than the permutations corresponding to symmetry groups.
We say that a Boolean function $f(x_1, x_2, \ldots, x_n)$ 
is \textbf{strongly asymmetric}
if for any permutation $\pi$ of $\{1, 2, \ldots, n\}$
\emph{except} the identity permutation,
there exists an input $(a_1, a_2, \ldots,  a_n)$ 
to $f$ such that $f(a_1, a_2, \ldots, a_n)$ $\neq$  
$f(a_{\pi(1)}, a_{\pi(2)}, \ldots, a_{\pi(n)})$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fi

The notion of strong asymmetry of Boolean functions was defined 
in Section~\ref{sec:prelim}.
In general, there are functions with $n$ variables that are properly
$n$-symmetric, but not strongly asymmetric \cite{KS-2000}.
We now show that any $n$ variable NCF that is properly $n$-symmetric
is also strongly asymmetric.
As a consequence, we point out there are NCFs with $n$ variables
which are properly $n$-symmetric.
